Embedding Dimension of Polyhedral Losses

### 2020-06-15 colt-20/embed-dim.tex
Clarifying notation in Sec 4

\begin{condition}[Halfspace condition]\label{cond:H-condition}
	A collection of polytopes $\T$ and a polytope $C\subseteq\simplex$ defined by $C = \{p \in \simplex : Bp \geq \vec 0\}$ \emph{satisfy the halfspace condition} if there exist $v_1, \ldots, v_k \in \reals^d$ such that, for all $i \in [k]$ and $y \in \Y$, for all $x \in T_y$, we have $\inprod{v_i}{x} \leq B_{iy}$.
\end{condition}
\begin{condition}[Vertex condition]\label{cond:V-condition}
	A collection of polytopes $\T$ and a polytope $C\subseteq\simplex$ defined by $C = \conv(\{p^1, \ldots, p^\ell\})$ \emph{satisfy the vertex condition} if for all $j \in [\ell]$, $0 \in \oplus_{p^j} \T$. %% Equivalent: there exist $\{x_{jy} \in T_y : j \in [\ell]\}$ such that, for all $j \in [\ell]$, $\sum_{y \in \Y} p^j_y x_{jy} = \vec 0$.
\end{condition}

\begin{theorem} \label{thm:vertex-halfspace-opt}
  Let the polytopes $\T = \{T_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
  We have $D(\T) = C$ if and only if both the halfspace and vertex conditions hold.
\end{theorem}

@@

Our program will consist of variables for the normal vectors $\{v_i \in \reals^d : i \in [k]\}$ for the (relaxed) halfspace condition, as described above, and variables for vertices $\{x^j_y \in \reals^d : j \in [\ell], y \in \Y\}$ which witness $0\in\oplus_{p^j} \T$ for the vertex condition, where the vector $x^j_y$ is the $y^{th}$ column of $X^j$.
% These correspond to a (possibly relaxed) halfspace representation and (possibly incomplete) vertex representation of the subgradient polytopes $\T$.

\begin{definition}[Quadratic Feasibility Program] \label{def:qfp} ~ \\
  \indent 
  \textbf{Given:} $d \in \mathbb{N}$, a polytope $C = \{p \in \simplex : Bp \geq \vec 0\} = \conv(\{p^1, \ldots, p^\ell\}) \subseteq \simplex$, where $B \in \reals^{k \times n}$ has a minimum number of rows.

  \textbf{Variables:} $V \in \reals^{k \times d}$ with rows $\{v_i\}$; $X^1,\ldots,X^{\ell} \in \reals^{d \times n}$, where $X^j$ has columns $\{x^j_y\}$.

@@

  ($\implies$)
  By the vertex condition, for each $j \in [\ell]$, there exist witnesses $\{x^j_y \in T_y : y \in \Y\}$ satisfying the second constraint of the quadratic program (Inequality \ref{eqn:qp-constr-2}).
  By the halfspace condition, there exist normals $v_1, \dots, v_k$ such that, for all $i$, for all $x \in T_y$, $\inprod{v_i}{x} \leq B_{iy}$; in particular, this applies to the above witnesses $x^j_y \in T_y$.
  Collecting $v_1,\dots,v_k$ as the columns of $V$, this shows that the first constraint (Inequality \ref{eqn:qp-constr-1}) is satisfied.

  ($\impliedby$)
  We construct $T_y = \conv(\{x^1_y, \ldots, x^{\ell}_y\})$.
  The second constraint of the quadratic program immediately implies the vertex condition.
  Taking $v_1,\dots,v_k$ as the columns of $V$, the first constraint implies that for each $x^j_y$, we have $\inprod{v_i}{x^j_y} \leq B_{iy}$ for all $i,j,y$.
  Any point $x \in T_y$ is a convex combination of $x^1_y,\ldots,x^{\ell}_y$, so it satisfies $\inprod{v_i}{x} \leq B_{iy}$.
  This implies the halfspace condition.  % Bo: QED
\end{proof}

>>>>

\begin{condition}[Halfspace condition]\label{cond:H-condition}
	A collection of polytopes $\T$ and a polytope $C\subseteq\simplex$ defined by $C = \{p \in \simplex : Bp \geq \vec 0\}$ \emph{satisfy the halfspace condition} if there exist $v_1, \ldots, v_k \in \reals^d$ such that, for all $i \in [k]$ and $y \in \Y$, for all $x \in T_y$, we have $\inprod{v_i}{x} \leq B_{iy}$.
\end{condition}
\begin{condition}[Vertex condition]\label{cond:V-condition}
	A collection of polytopes $\T$ and a polytope $C\subseteq\simplex$ defined by $C = \conv(\{p^1, \ldots, p^l\})$ \emph{satisfy the vertex condition} if for all $j \in [l]$, $0 \in \oplus_{p^j} \T$. %% Equivalent: there exist $\{x_{jy} \in T_y : j \in [\ell]\}$ such that, for all $j \in [\ell]$, $\sum_{y \in \Y} p^j_y x_{jy} = \vec 0$.
\end{condition}

\begin{theorem} \label{thm:vertex-halfspace-opt}
  Let the polytopes $\T = \{T_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{l}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
  We have $D(\T) = C$ if and only if both the halfspace and vertex conditions hold.
\end{theorem}

@@

Our program will consist of variables for the normal vectors $\{v_i \in \reals^d : i \in [k]\}$ for the (relaxed) halfspace condition, as described above, and variables for vertices $\{x^j_y \in \reals^d : j \in [l], y \in \Y\}$ which witness $0\in\oplus_{p^j} \T$ for the vertex condition, where the vector $x^j_y$ is the $y^{th}$ column of $X^j$.
% These correspond to a (possibly relaxed) halfspace representation and (possibly incomplete) vertex representation of the subgradient polytopes $\T$.

\begin{definition}[Quadratic Feasibility Program] \label{def:qfp} ~ \\
  \indent 
  \textbf{Given:} $d \in \mathbb{N}$, a polytope $C = \{p \in \simplex : Bp \geq \vec 0\} = \conv(\{p^1, \ldots, p^l\}) \subseteq \simplex$, where $B \in \reals^{k \times n}$ has a minimum number of rows.

  \textbf{Variables:} $V \in \reals^{k \times d}$ with rows $\{v_i\}$; $X^1,\ldots,X^{l} \in \reals^{d \times n}$, where $X^j$ has columns $\{x^j_y\}$.

@@

  ($\implies$)
  By the vertex condition, for each $j \in [l]$, there exist witnesses $\{x^j_y \in T_y : y \in \Y\}$ satisfying the second constraint of the quadratic program (Inequality \ref{eqn:qp-constr-2}).
  By the halfspace condition, there exist normals $v_1, \dots, v_k$ such that, for all $i$, for all $x \in T_y$, $\inprod{v_i}{x} \leq B_{iy}$; in particular, this applies to the above witnesses $x^j_y \in T_y$.
  Collecting $v_1,\dots,v_k$ as the columns of $V$, this shows that the first constraint (Inequality \ref{eqn:qp-constr-1}) is satisfied.

  ($\impliedby$)
  We construct $T_y = \conv(\{x^1_y, \ldots, x^{l}_y\})$.
  The second constraint of the quadratic program immediately implies the vertex condition.
  Taking $v_1,\dots,v_k$ as the columns of $V$, the first constraint implies that for each $x^j_y$, we have $\inprod{v_i}{x^j_y} \leq B_{iy}$ for all $i,j,y$.
  Any point $x \in T_y$ is a convex combination of $x^1_y,\ldots,x^{l}_y$, so it satisfies $\inprod{v_i}{x} \leq B_{iy}$.
  This implies the halfspace condition.  % Bo: QED
\end{proof}
